Malacology: A Programmable Storage System Built on Ceph

### 2016-05-08 paper/paper.tex
paper: section 2
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

A consequence of complicated data management frameworks and faster
devices is that the storage system cannot meet the needs of the
general-purpose storage system. Workarounds for helping the application
meet its performance goals roughly fall into one of three categories:
``bolt-on'' services, application changes, and storage changes.

@@

So called ``bolt-on'' services are 3rd party systems that are integrated
into the system to provide functionality and improve performance for
common use-cases (e.g.~a metadata service). In addition to sacrificing
simplicity, ``bolt-on'' services often duplicate functionality and
execute redundant code, unnecessarily increasing the likelihood of bugs
or worse, introducing unpredictable performance problems. For example,
we are developing a distributed shared-commit log on Ceph called ZLog.
ZLog uses a ``sequencer'' to distribute tokens to clients that want to
append to the end of the log. When designing the sequencer, we
considered bolting on Zookeeper \autocite{} as a service. Zookeeper would
satisfy our requirements by proiding a network and transaction stack but
it would perform poorly because Zookeeper persists its data and the
sequencer can be in-memory and volatile. A Zookeeper sequencer would
adhere to the pre-conceived notions of storage, namely that data in
storage systems should be safe from data loss, but would introduce
unecessary scalability limits \autocite{Neha HDFS namenode example}.

@@

The second approach to adapting to a storage system deficiency is to
extend the responsibility of the application. This means changing the
application itself by adding more data management intelligence or
domain-specific middleware (e.g., a new data layout). For instance, an
application may change itself to exploit data locality or I/O
parallelism in a distributed storage system.

For example, SciHadoop \autocites[
]{buck:hpc2011-scihadoop}{buck:sc2013-scidr} changes both the Hadoop
application and the Hadoop framework itself to leverage the structure of
scientific (3D array-based data, more specifically) to increase
performance, locality, and the number of early results. This is not a
bad proposition, but creates a coupling that is highly tied to the
underlying physical properties of the system, making it difficult to
adapt to future changes at the storage system level.

\subsection{Storage Changes}\label{storage-changes}

@@

Tuning large and complicated systems is difficult because of the number
of parameters (e.g., Hadoop 2.7.1 has 210 tunables) and because the
tunables are difficult to understand or quantify (e.g., Ceph tunables
\autocite{sevilla:sc15-mantle}). To succesfully tune the system, the
developer must have domain and system specific knowledge. Without this
intimate knowledge, the tuning turns into somewhat of a black art, where
the only way to figure out the best settings is trial and error.

Auto-tuning techniques attempt to find a good solution among a huge
space of available system configurations. However, in practice
auto-tuning is limited to only the configuration tunables that the
storage system exposes (e.g.~block size) and can be overwhelmed with too
many parameters. Starfish \autocite{herodotou:cidr2011-starfish} made an
attempt of this for Hadoop but this technique would need to severely
limit the space of parameters in order to feasible, similar to the
approach used in \autocite{behzad:sc2013-autotuning}. For instance,
auto-tuning may be capable of identifying instances in which new data
layouts would benefit a workload, but unless the system can provide such
a transformation, the option is left off the table.

@@

As a last resort, the application developer can introduce changes to the
storage system itself. This endeavour is especially difficult because
the developer must first familiarate themselves with the storage system,
and even if they manage to make the necessary changes, they must get
their changes upstream for that specific project or become mantainer for
their in-house version. Being a maintainer means they re-base their
versions against new versions of the storage system and address any bugs
for their branched code. A successful example of this is the work that
focused on HDFS scalability for metadata-intensive workloads
\autocite{shvachko_hdfs_2010}. This has lead to modifications to its
architecture or API \autocite{balmin:sigmod2012-clydesdale} to improve
performance.

\begin{table}[ht]
\caption{
    A variety of RADOS object storage classes exist to expose interfaces
    to applications. \# is the number of methods that use these categories.
    \label{table:objclasses}}
  \begin{center}
  \begin{tabular}{l|l|l}
    Category & Specialization& \# \\ \hline
    \multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
                             & Exclusive & \\ \hdashline
    \multirow{3}{*}{Logging} & Replica & 3 \\
                             & State & 4 \\
                             & Timestamped & 4 \\ \hdashline
    \multirow{4}{*}{Metadata Managment} 
                             & RADOS Block Device  & 37 \\
                             & RADOS Gatway & 27 \\
                             & User & 5 \\
                             & Version & 5 \\ \hdashline
    Garbage Collection       & Reference Counting & 4 \\
\end{tabular}
\end{center}
\end{table}

\subsubsection{Hybrid Approaches}\label{hybrid-approaches}

As an alternative, some systems allow developers to extend the
interfaces of the storage devices to perform arbitrary operations on
data stored on disk. In Ceph, object interfaces are C/C++ plugins that
add new functionality to the OSDs. As an example use-case, ZLog needs to
store metadata information with each object so for each operation, it
checks the epoch value and write capability (2 metadata reads) and
writes the index (metadata write). Figure 1 shows the throughput
(\emph{y} axis) over time (\emph{x} axis) of two OSD implementations for
storing metadata: in the header of the byte stream (data) vs.~in the
object extended attributes (XATTR). The speed for appending data without
any metadata operations (data raw) is also shown as a baseline for
comparison. For append-heavy workloads storing metadata as a header in
the data performs about 1.5x better than storing metadata as an extended
attribute.

@@

Object interface classese are used in production Ceph environments;
Figure \ref{fig:obj-int-dev-growth} shows a dramatic growth in the use
of co-designed interfaces in the Ceph community since 2010. Figure
\ref{fig:obj-int-dev-churn} examines this growth in interfaces further
by showing the lines of code that are changed (\emph{y} axis) over time
(\emph{x} axis). The prevalence of blue dots indicates that users
frequently update their interfaces close to release cycles and that
massive changes come in bunches. This interface churn reflects both the
popularity of object interfaces and the complexity of the processing
being done by the OSDs. What is most remarkable about Figure
\ref{fig:obj-int-dev-churn} is that this trend contradicts the notion
that API changes are a burden for users. The types of object interfaces,
which are enumerated in Table \ref{table:objclasses}, show that this
active storage development is present throughough the Ceph software
stack.

\begin{figure}[htbp]
\centering
\includegraphics{figures/obj-int-dev-growth.png}
\caption{Since 2010, the growth in the number of co-designed object
storage interfaces in Ceph has been accelerating. This plot is the
number of object classes (a group of interfaces), and the total number
of methods (the actual API end-points).\label{fig:obj-int-dev-growth}}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figures/obj-int-dev-churn.png}
\caption{Source code changes over time indicate the dynamic interface
development will need to support a high frequency of interface churn.
Each dot represents a Ceph commit with a corresponding number of lines
of code changed. \label{fig:obj-int-dev-churn}}
\end{figure}

@@

Ceph provides storage by striping and replicating data across RADOS
\autocite{weil_rados_2007}, the reliable distributed object store. RADOS
uses many techniques to ensure that data is not corrupted or lost, such
as erasure coding, replication, and data scrubbing. Furthermore, many of
these techniques try to be autonomous so that work is distributed across
the cluster. For example, when placement groups change the OSDs
rebalance and re-shard data in the background in a process called
placement group splitting.

>>>>

A consequence of complicated data management requirements and faster storage
devices is that storage systems may fail to meet the needs of applications.
Application workarounds for meeting performance goals roughly fall into one of
three categories: ``bolt-on'' services, application changes, and changes to
the storage system itself.

@@

So called ``bolt-on'' services are third-party solutions integrated into a
system to provide functionality and improve performance for common use-cases
(e.g. a metadata service). The introduction of additional services may provide
simplicity and performance enhancements, but ``bolt-on`` services often
duplicate functionality, unnecessarily increasing the likelihood of bugs or
introducing unpredictable performance problems. For example, we are developing
a high-performance distributed shared-log on Ceph called ZLog, based on the
CORFU protocol~\cite{corfu}. Serialization at high-performance is achieved using a
``sequencer'' service that maintains a volatile, in-memory counter used to assign a log
position to each client that appends to the log. One option for building
a distributed log is to use PAXOS, or an out-of-the-box solution such as
Zookeeper. However, these systems adhere to the pre-coneived notions of
storage, namely that data is safe from data loss, which is exactly the feature
that causes traditional approaches to run at significantly reduced speed.

@@

The second approach that applications may use to adapt to a storage system
limitation is to
extend the responsibility of the application. This means changing the
application itself by adding more data management intelligence or
using domain-specific middleware (e.g., a new data layout). For instance, an
application may be changed to exploit data locality or I/O
parallelism in a distributed storage system.

For example, SciHadoop \autocites[
]{buck:hpc2011-scihadoop}{buck:sc2013-scidr} changes both the Hadoop
application and the Hadoop framework itself to leverage the structure of
scientific data (3D array-based data, more specifically) to increase
performance, locality, and the number of early results. This is not a
bad proposition, but creates a coupling that is highly tied to the
underlying physical properties of the system, making it difficult to
adapt to future changes at the storage system level.

\subsection{Storage System Changes}\label{storage-changes}

@@

Tuning large and complicated systems is difficult because of the number
of parameters (e.g., Hadoop 2.7.1 has 210 tunables) and because the
tunables are difficult to understand or quantify (e.g., Ceph tunables
\autocite{sevilla:sc15-mantle}). To succesfully tune the system, a
developer must have domain and system specific knowledge. Without this
intimate knowledge, the tuning turns into somewhat of a black art, where
the only way to figure out the best settings is trial and error.

Auto-tuning techniques attempt to find a good solution among a huge space of
available system configurations. However, in practice auto-tuning is limited
to only the configuration tunables that the storage system exposes and can be
overwhelmed with too many parameters.  For instance, auto-tuning may be
capable of identifying instances in which new data layouts would benefit a
workload, but unless the system can provide such a transformation, the option
is left off the table.  Starfish \autocite{herodotou:cidr2011-starfish} made
an attempt of this for Hadoop but this technique would need to severely limit
the space of parameters in order to feasible, similar to the approach used in
\autocite{behzad:sc2013-autotuning}.

@@

As a last resort, the application developer can alter the storage system
itself. In the best case a developer is using an open-source system in which
case one must become familiar with often enormous code-bases before being able
to make changes, but then face the task of approaching a community to
integrate changes, or be forced to maintain an in-house version and risk
maintaining features that become incompatible.  A successful example of this
is the work that focused on HDFS scalability for metadata-intensive workloads
\autocite{shvachko_hdfs_2010}. This has lead to modifications to its
architecture or API \autocite{balmin:sigmod2012-clydesdale} to improve
performance.

Alternatively, some systems have formal support for introducing additional
functionality, often in the form of a plugin infrastructure. Ceph provides the
ability to create domain-specific logical object interfaces by composing
existing interfaces. These interfaces are written in C++ and are statically
loaded into the system. An example usage may be to atomically combine a
``write'' with an index update to create a content-addressable object
interface.

\being{comment}
As an example use-case, ZLog needs to
store metadata information with each object so for each operation, it
checks the epoch value and write capability (2 metadata reads) and
writes the index (metadata write).
\end{comment}

Figure 1 shows the throughput
(\emph{y} axis) over time (\emph{x} axis) of two OSD implementations for
storing metadata: in the header of the byte stream (data) vs.~in the
object extended attributes (XATTR). The speed for appending data without
any metadata operations (data raw) is also shown as a baseline for
comparison. For append-heavy workloads storing metadata as a header in
the data performs about 1.5x better than storing metadata as an extended
attribute.

@@

The object interface facility provided by Ceph is more than an
proof-of-concept active storage implementation.  Figure
\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
domain-specific interfaces in the Ceph community since 2010.  What is most
remarkable is that this trend contradicts the notion that API changes are a
burden for users. Rather it appears that a gap in existing interfaces are
being addressed through programmability. In fact, Table~\ref{table:objclasses}
categorizes existing interfaces and we clearly see a trend towards reusable
services.

\begin{figure}[ht]
\centering
\includegraphics{figures/obj-int-dev-growth.png}
\caption{Since 2010, the growth in the number of co-designed object
storage interfaces in Ceph has been accelerating. This plot is the
number of object classes (a group of interfaces), and the total number
of methods (the actual API end-points).\label{fig:obj-int-dev-growth}}
\end{figure}

\begin{table}[ht]
\centering
  \begin{tabular}{l|l|l}
    Category & Specialization& \# \\ \hline
    \multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
                             & Exclusive & \\ \hdashline
    \multirow{3}{*}{Logging} & Replica & 3 \\
                             & State & 4 \\
                             & Timestamped & 4 \\ \hdashline
    \multirow{4}{*}{Metadata Managment} 
                             & RADOS Block Device  & 37 \\
                             & RADOS Gatway & 27 \\
                             & User & 5 \\
                             & Version & 5 \\ \hdashline
    Garbage Collection       & Reference Counting & 4 \\
\end{tabular}
\caption{A variety of RADOS object storage classes exist to expose interfaces
    to applications. \# is the number of methods that use these categories.
}
\label{table:objclasses}
\end{table}

\begin{comment}
\begin{figure}[htbp]
\centering
\includegraphics{figures/obj-int-dev-churn.png}
\caption{Source code changes over time indicate the dynamic interface
development will need to support a high frequency of interface churn.
Each dot represents a Ceph commit with a corresponding number of lines
of code changed. \label{fig:obj-int-dev-churn}}
\end{figure}
\end{comment}

@@

Ceph provides storage by striping and replicating data across RADOS
~\cite{weil_rados_2007}, the reliable distributed object store. RADOS
uses many techniques to ensure that data is not corrupted or lost, such
as erasure coding, replication, and data scrubbing. Furthermore, many of
these techniques try to be autonomous so that work is distributed across
the cluster. For example, when placement groups change the OSDs
rebalance and re-shard data in the background in a process called
placement group splitting.
